---
title: "Final Project Notebook"
output: html_notebook
---

```{r}
rm(list = ls())
library(tidyverse)
```

Read in dataset from previous project
```{r}
df = read_delim("drugstudy.csv", delim = ',') %>% subset(select = -ID)
```

Setup individual datasets
```{r}
m1 = df %>% select(Gender, Age, Race, Employment, Cigarette, Theft, Burglary, Arson) %>% 
  mutate("CrimeYesNo" = ifelse((Theft == "Yes") | (Burglary == "Yes") | (Arson == "Yes"), 1, 0)) %>%
  subset(select = -c(Theft, Burglary, Arson))
m1 = m1[complete.cases(m1$CrimeYesNo),]
m1$Cigarette = ifelse(m1$Cigarette == "Yes", 1, 0)

m2 = df %>% select(Gender, Age, Race, Employment, Marijuana, Theft, Burglary, Arson) %>% 
  mutate("CrimeYesNo" = ifelse((Theft == "Yes") | (Burglary == "Yes") | (Arson == "Yes"), 1, 0)) %>%
  subset(select = -c(Theft, Burglary, Arson))
m2 = m2[complete.cases(m2$CrimeYesNo),]
m2$Marijuana = ifelse(m2$Marijuana == "Yes", 1, 0)

worst_feelings = c("NervousWorst", "HopelessWorst", "WorthlessWorst")
m3 = df %>% select(Gender, Age, Race, Employment, Cigarette, NervousWorst, HopelessWorst, WorthlessWorst)
m3[worst_feelings] = m3[worst_feelings] %>% 
  mutate_all(. ~recode(., "None of the time" = 0, 
                       "A little of the time" = 1, "Some of the time" = 3, 
                       "Most of the time" = 4, "All of the time" = 5))
m3 = m3 %>% 
  mutate("feelings" = ifelse((NervousWorst > 2) | (HopelessWorst > 2) | (WorthlessWorst > 2), 1, 0),
         "Cigarette" = ifelse(Cigarette == "Yes", 1, 0)) %>%  
  subset(select = -c(NervousWorst, HopelessWorst, WorthlessWorst))
m3 = m3[complete.cases(m3$feelings),]

m4 = df %>% select(Gender, Age, Race, Employment, Marijuana, NervousWorst, HopelessWorst, WorthlessWorst)
m4[worst_feelings] = m4[worst_feelings] %>% 
  mutate_all(. ~recode(., "None of the time" = 0, 
                       "A little of the time" = 1, "Some of the time" = 3, 
                       "Most of the time" = 4, "All of the time" = 5))
m4 = m4 %>% 
  mutate("feelings" = ifelse((NervousWorst > 2) | (HopelessWorst > 2) | (WorthlessWorst > 2), 1, 0),
         "Marijuana" = ifelse(Marijuana == "Yes", 1, 0)) %>%  
  subset(select = -c(NervousWorst, HopelessWorst, WorthlessWorst))
m4 = m4[complete.cases(m4$feelings),]
```


CEM - Coarsened Exact Matching
```{r}
library(caret)
library(cem)
covs = c("Gender", "Age", "Race", "Employment")
```

Model 1: Cigarette / Crimes
```{r}
imbalance(group=m1$Cigarette, 
          data= as.data.frame(predict(dummyVars(~ Gender + Age + Race + Employment, m1[covs]), m1[covs])))
```

```{r}
m1_vars = as.data.frame(cbind(predict(dummyVars(~ Gender + Age + Race + Employment, m1[covs]), m1[covs]), 
                              "Cigarette" = m1$Cigarette, "CrimeYesNo" = m1$CrimeYesNo))
cem_1 = cem(treatment = "Cigarette", data = m1_vars, drop = "CrimeYesNo")
att_1 = att(cem_1, CrimeYesNo ~ Cigarette, data = m1_vars, model = "logistic")
summary(att_1)

# Extrapolation
att_1_extra <- att(cem_1, CrimeYesNo ~ Cigarette, data = m1_vars, model="logistic", extra=TRUE)
summary(att_1_extra)
```

Interpretation
```{r}
(exp(0.29464) - 1) * 100
(exp(0.22347) - 1) * 100
```

```{r}
plot(att_1, cem_1, m1_vars)
```


Model 2: Marijuana / Crimes
```{r}
imbalance(group=m2$Marijuana,
          data= as.data.frame(predict(dummyVars(~ Gender + Age + Race + Employment, m2[covs]), m2[covs])))
```

```{r}
m2_vars = as.data.frame(cbind(predict(dummyVars(~ Gender + Age + Race + Employment, m2[covs]), m2[covs]), 
                              "Marijuana" = m2$Marijuana, "CrimeYesNo" = m2$CrimeYesNo))
cem_2 = cem(treatment = "Marijuana", data = m2_vars, drop = "CrimeYesNo")
att_2 = att(cem_2, CrimeYesNo ~ Marijuana, data = m2_vars, model = "logistic")
summary(att_2)

# Extrapolation
att_2_extra <- att(cem_2, CrimeYesNo ~ Marijuana, data = m2_vars, model="logistic", extra=TRUE)
summary(att_2_extra)
```

Interpretation
```{r}
(exp(0.10874) - 1) * 100
(exp(0.084459) - 1) * 100
```

```{r}
plot(att_2, cem_2, m2_vars)
```


Model 3: Cigarette / Stress
```{r}
imbalance(group=m3$Cigarette, 
          data= as.data.frame(predict(dummyVars(~ Gender + Age + Race + Employment, m3[covs]), m3[covs])))
```

```{r}
m3_vars = as.data.frame(cbind(predict(dummyVars(~ Gender + Age + Race + Employment, m3[covs]), m3[covs]), 
                              "Cigarette" = m3$Cigarette, "feelings" = m3$feelings))
cem_3 = cem(treatment = "Cigarette", data = m3_vars, drop = "feelings")
att_3 = att(cem_3, feelings ~ Cigarette, data = m3_vars, model = "logistic")
summary(att_3)

# Extrapolation
att_3_extra <- att(cem_3, feelings ~ Cigarette, data = m3_vars, model="logistic", extra=TRUE)
summary(att_3_extra)
```

Interpretation
```{r}
(exp(0.224117) - 1) * 100
(exp(0.222468) - 1) * 100
```

```{r}
plot(att_3, cem_3, m3_vars)
```


Model 4: Marijuana / Stress
```{r}
imbalance(group=m4$Marijuana,
          data= as.data.frame(predict(dummyVars(~ Gender + Age + Race + Employment, m4[covs]), m4[covs])))
```

```{r}
m4_vars = as.data.frame(cbind(predict(dummyVars(~ Gender + Age + Race + Employment, m4[covs]), m4[covs]), 
                              "Marijuana" = m4$Marijuana, "feelings" = m4$feelings))
cem_4 = cem(treatment = "Marijuana", data = m4_vars, drop = "feelings")
att_4 = att(cem_4, feelings ~ Marijuana, data = m4_vars, model = "logistic")
summary(att_4)

# Extrapolation
att_4_extra <- att(cem_4, feelings ~ Marijuana, data = m4_vars, model="logistic", extra=TRUE)
summary(att_4_extra)
```

Interpretation
```{r}
(exp(0.391917) - 1) * 100
(exp(0.389258) - 1) * 100
```

```{r}
plot(att_4, cem_4, m4_vars)
```


Doubly Robust Regression - Bootstrap estimates and then create confidence interval for ATE

Model 1: Cigarette/Crimes
```{r, warning=FALSE}
att_1 = c()
for(i in 1:1000){
  if(i %% 100 == 0){print(paste("Iteration", i, "complete"))}
  indices = sample(1:nrow(m1), size = 200, replace = TRUE)
  temp = m1[indices,]
  trt = temp$Cigarette
  y = temp$CrimeYesNo
  out_mod = glm(CrimeYesNo ~ Cigarette + Gender + Age + Race + Employment, data = temp, family = "binomial")
  ps_mod = glm(Cigarette ~ Gender + Age + Race + Employment, data = temp, family = "binomial")
  e_hat = ps_mod$fitted.values
  dat_1 = transform(temp, Cigarette = 1)
  dat_0 = transform(temp, Cigarette = 0)
  y_hat_1 = predict(out_mod, newdata = dat_1, type = "response") 
  y_hat_0 = predict(out_mod, newdata = dat_0, type = "response")
  DR_1 = (e_hat^-1)*y*trt - (e_hat^-1)*(y_hat_1*(trt - e_hat))
  DR_0 = (1 - e_hat)^-1*(y*(1 - trt)) + (1 - e_hat)^-1*(y_hat_0*(trt - e_hat))
  att_1 = c(att_1, mean(DR_1) - mean(DR_0))
}

mean_att = mean(att_1)
sd_att = sqrt(var(att_1))/sqrt(length(att_1))
print("Confidence Interval:")
c(mean_att - 1.96*sd_att, mean_att + 1.96*sd_att)
```

Model 2: Marijuana/Crimes
```{r, warning=FALSE}
att_2 = c()
for(i in 1:1000){
  if(i %% 100 == 0){print(paste("Iteration", i, "complete"))}
  indices = sample(1:nrow(m2), size = 200, replace = TRUE)
  temp = m2[indices,]
  trt = temp$Marijuana
  y = temp$CrimeYesNo
  out_mod = glm(CrimeYesNo ~ Marijuana + Gender + Age + Race + Employment, data = temp, family = "binomial")
  ps_mod = glm(Marijuana ~ Gender + Age + Race + Employment, data = temp, family = "binomial")
  e_hat = ps_mod$fitted.values
  dat_1 = transform(temp, Marijuana = 1)
  dat_0 = transform(temp, Marijuana = 0)
  y_hat_1 = predict(out_mod, newdata = dat_1, type = "response") 
  y_hat_0 = predict(out_mod, newdata = dat_0, type = "response")
  DR_1 = (e_hat^-1)*y*trt - (e_hat^-1)*(y_hat_1*(trt - e_hat))
  DR_0 = (1 - e_hat)^-1*(y*(1 - trt)) + (1 - e_hat)^-1*(y_hat_0*(trt - e_hat))
  att_2 = c(att_2, mean(DR_1) - mean(DR_0))
}

mean_att = mean(att_2)
sd_att = sqrt(var(att_2))/sqrt(length(att_2))
print("Confidence Interval:")
c(mean_att - 1.96*sd_att, mean_att + 1.96*sd_att)
```

Model 3: Cigarette/Emotional Stress
```{r, warning=FALSE}
att_3 = c()
for(i in 1:1000){
  if(i %% 100 == 0){print(paste("Iteration", i, "complete"))}
  indices = sample(1:nrow(m3), size = 200, replace = TRUE)
  temp = m3[indices,]
  trt = temp$Cigarette
  y = temp$feelings
  out_mod = glm(feelings ~ Cigarette + Gender + Age + Race + Employment, data = temp, family = "binomial")
  ps_mod = glm(Cigarette ~ Gender + Age + Race + Employment, data = temp, family = "binomial")
  e_hat = ps_mod$fitted.values
  dat_1 = transform(temp, Cigarette = 1)
  dat_0 = transform(temp, Cigarette = 0)
  y_hat_1 = predict(out_mod, newdata = dat_1, type = "response") 
  y_hat_0 = predict(out_mod, newdata = dat_0, type = "response")
  DR_1 = (e_hat^-1)*y*trt - (e_hat^-1)*(y_hat_1*(trt - e_hat))
  DR_0 = (1 - e_hat)^-1*(y*(1 - trt)) + (1 - e_hat)^-1*(y_hat_0*(trt - e_hat))
  att_3 = c(att_3, mean(DR_1) - mean(DR_0))
}

mean_att = mean(att_3)
sd_att = sqrt(var(att_3))/sqrt(length(att_3))
print("Confidence Interval:")
c(mean_att - 1.96*sd_att, mean_att + 1.96*sd_att)
```

Model 4: Marijuana/Emotional Stress
```{r, warning=FALSE}
att_4 = c()
for(i in 1:1000){
  if(i %% 100 == 0){print(paste("Iteration", i, "complete"))}
  indices = sample(1:nrow(m2), size = 200, replace = TRUE)
  temp = m4[indices,]
  trt = temp$Marijuana
  y = temp$feelings
  out_mod = glm(feelings ~ Marijuana + Gender + Age + Race + Employment, data = temp, family = "binomial")
  ps_mod = glm(Marijuana ~ Gender + Age + Race + Employment, data = temp, family = "binomial")
  e_hat = ps_mod$fitted.values
  dat_1 = transform(temp, Marijuana = 1)
  dat_0 = transform(temp, Marijuana = 0)
  y_hat_1 = predict(out_mod, newdata = dat_1, type = "response") 
  y_hat_0 = predict(out_mod, newdata = dat_0, type = "response")
  DR_1 = (e_hat^-1)*y*trt - (e_hat^-1)*(y_hat_1*(trt - e_hat))
  DR_0 = (1 - e_hat)^-1*(y*(1 - trt)) + (1 - e_hat)^-1*(y_hat_0*(trt - e_hat))
  att_4 = c(att_4, mean(DR_1) - mean(DR_0))
}

mean_att = mean(att_4)
sd_att = sqrt(var(att_4))/sqrt(length(att_4))
print("Confidence Interval:")
c(mean_att - 1.96*sd_att, mean_att + 1.96*sd_att)
```


BART - Bayesian Additive Regression Tree

```{r}
library(caret)
library(BayesTree)
```

Model 1: Cigarette / Crimes
```{r}
bart1 = m1
bart1$Age = as.character(bart1$Age)
bart1_ohe_vars = predict(dummyVars(~ Gender + Age + Race + Employment, bart1), bart1)
bart1 = as.data.frame(cbind(bart1_ohe_vars, 
                            "Cigarette" = bart1$Cigarette, 
                            "CrimeYesNo" = bart1$CrimeYesNo)) %>%
  arrange(desc(CrimeYesNo))
Xtrain = bart1[, !colnames(bart1) %in% "CrimeYesNo"]
Y = bart1$CrimeYesNo
xp1 <- xp2 <- bart1[bart1$Cigarette == 1, !colnames(bart1) %in% "CrimeYesNo"]
xp2$Cigarette = 0
Xtest = rbind(xp1, xp2)
model_1 = bart(x.train = Xtrain, y.train = Y, x.test = Xtest)
nt = ncol(model_1$yhat.test)/2
tmp = apply(model_1$yhat.test[,1:nt] -
              model_1$yhat.test[,(nt+1):(2*nt)], 1, mean)
mean_att = mean(tmp)
sd_att = sqrt(var(tmp)) / length(tmp)
print("Confidence Interval:")
c(mean_att - 1.96*sd_att, mean_att + 1.96*sd_att)
```

Model 2: Marijuana / Crimes
```{r}
bart2 = m2
bart2$Age = as.character(bart2$Age)
bart2_ohe_vars = predict(dummyVars(~ Gender + Age + Race + Employment, bart2), bart2)
bart2 = as.data.frame(cbind(bart2_ohe_vars, 
                            "Marijuana" = bart2$Marijuana, 
                            "CrimeYesNo" = bart2$CrimeYesNo)) %>%
  arrange(desc(CrimeYesNo))
Xtrain = bart2[, !colnames(bart2) %in% "CrimeYesNo"]
Y = bart2$CrimeYesNo
xp1 <- xp2 <- bart2[bart2$Marijuana == 1, !colnames(bart2) %in% "CrimeYesNo"]
xp2$Marijuana = 0
Xtest = rbind(xp1, xp2)
model_2 = bart(x.train = Xtrain, y.train = Y, x.test = Xtest)
nt = ncol(model_2$yhat.test)/2
tmp = apply(model_2$yhat.test[,1:nt] -
              model_2$yhat.test[, (nt+1):(2*nt)], 1, mean)
mean_att = mean(tmp)
sd_att = sqrt(var(tmp)) / length(tmp)
print("Confidence Interval:")
c(mean_att - 1.96*sd_att, mean_att + 1.96*sd_att)
```

Model 3: Cigarette / Stress
```{r}
bart3 = m3
bart3$Age = as.character(bart3$Age)
bart3_ohe_vars = predict(dummyVars(~ Gender + Age + Race + Employment, bart3), bart3)
bart3 = as.data.frame(cbind(bart3_ohe_vars, 
                            "Cigarette" = bart3$Cigarette, 
                            "feelings" = bart3$feelings)) %>%
  arrange(desc(feelings))
Xtrain = bart3[, !colnames(bart3) %in% "feelings"]
Y = bart3$feelings
xp1 <- xp2 <- bart3[bart3$Cigarette == 1, !colnames(bart3) %in% "feelings"]
xp2$Cigarette = 0
Xtest = rbind(xp1, xp2)
model_3 = bart(x.train = Xtrain, y.train = Y, x.test = Xtest)
nt = ncol(model_3$yhat.test)/2
tmp = apply(model_3$yhat.test[,1:nt] -
              model_3$yhat.test[,(nt+1):(2*nt)], 1, mean)
mean_att = mean(tmp)
sd_att = sqrt(var(tmp)) / length(tmp)
print("Confidence Interval:")
c(mean_att - 1.96*sd_att, mean_att + 1.96*sd_att)
```

Model 4: Marijuana / Stress
```{r}
bart4 = m4
bart4$Age = as.character(bart4$Age)
bart4_ohe_vars = predict(dummyVars(~ Gender + Age + Race + Employment, bart4), bart4)
bart4 = as.data.frame(cbind(bart4_ohe_vars, 
                            "Marijuana" = bart4$Marijuana, 
                            "feelings" = bart4$feelings)) %>%
  arrange(desc(feelings))
Xtrain = bart4[, !colnames(bart4) %in% "feelings"]
Y = bart4$feelings
xp1 <- xp2 <- bart4[bart4$Marijuana == 1, !colnames(bart4) %in% "feelings"]
xp2$Marijuana = 0
Xtest = rbind(xp1, xp2)
model_4 = bart(x.train = Xtrain, y.train = Y, x.test = Xtest)
nt = ncol(model_4$yhat.test)/2
tmp = apply(model_4$yhat.test[,1:nt] -
              model_4$yhat.test[, (nt+1):(2*nt)], 1, mean)
mean_att = mean(tmp)
sd_att = sqrt(var(tmp)) / length(tmp)
print("Confidence Interval:")
c(mean_att - 1.96*sd_att, mean_att + 1.96*sd_att)
```


LEFTOVER WORK
For robust regression: 

Model 1:
```{r, eval=FALSE}
trt = m1$Cigarette
y = m1$CrimeYesNo
out_mod <- glm(CrimeYesNo ~ Cigarette + Gender + Age + Race + Employment, data = m1, family = "binomial")
ps_mod <- glm(Cigarette ~ Gender + Age + Race + Employment, data = m1, family = "binomial")
e_hat <- ps_mod$fitted.values
dat_1 <- transform(m1, Cigarette = 1)
dat_0 <- transform(m1, Cigarette = 0)
y_hat_1 <- predict(out_mod, newdata = dat_1, type = "response") 
y_hat_0 <- predict(out_mod, newdata = dat_0, type = "response")
boxplot(y_hat_1, y_hat_0)
DR_1 <- (e_hat^-1)*y*trt - (e_hat^-1)*(y_hat_1*(trt - e_hat))
DR_0 <- (1 - e_hat)^-1*(y*(1 - trt)) + (1 - e_hat)^-1*(y_hat_0*(trt - e_hat))
paste("Estimated treatment effect (ATE): ",mean(DR_1) - mean(DR_0))
```

```{r, eval=FALSE}
trt = m2$Marijuana
y = m2$CrimeYesNo
out_mod <- glm(CrimeYesNo ~ Marijuana + Gender + Age + Race + Employment, data = m2, family = "binomial")
ps_mod <- glm(Marijuana ~ Gender + Age + Race + Employment, data = m2, family = "binomial")
e_hat <- ps_mod$fitted.values
dat_1 <- transform(m2, Marijuana = 1)
dat_0 <- transform(m2, Marijuana = 0)
y_hat_1 <- predict(out_mod, newdata = dat_1, type = "response") 
y_hat_0 <- predict(out_mod, newdata = dat_0, type = "response")
boxplot(y_hat_1, y_hat_0)
DR_1 <- (e_hat^-1)*y*trt - (e_hat^-1)*(y_hat_1*(trt - e_hat))
DR_0 <- (1 - e_hat)^-1*(y*(1 - trt)) + (1 - e_hat)^-1*(y_hat_0*(trt - e_hat))
paste("Estimated treatment effect (ATE): ",mean(DR_1) - mean(DR_0))
```

```{r,eval=FALSE}
trt = m3$Cigarette
y = m3$feelings
out_mod <- glm(feelings ~ Cigarette + Gender + Age + Race + Employment, data = m3, family = "binomial")
ps_mod <- glm(Cigarette ~ Gender + Age + Race + Employment, data = m3, family = "binomial")
e_hat <- ps_mod$fitted.values
dat_1 <- transform(m3, Cigarette = 1)
dat_0 <- transform(m3, Cigarette = 0)
y_hat_1 <- predict(out_mod, newdata = dat_1, type = "response") 
y_hat_0 <- predict(out_mod, newdata = dat_0, type = "response")
boxplot(y_hat_1, y_hat_0)
DR_1 <- (e_hat^-1)*y*trt - (e_hat^-1)*(y_hat_1*(trt - e_hat))
DR_0 <- (1 - e_hat)^-1*(y*(1 - trt)) + (1 - e_hat)^-1*(y_hat_0*(trt - e_hat))
paste("Estimated treatment effect (ATE): ",mean(DR_1) - mean(DR_0))
```

```{r,eval=FALSE}
trt = m4$Marijuana
y = m4$feelings
out_mod <- glm(feelings ~ Marijuana + Gender + Age + Race + Employment, data = m4, family = "binomial")
ps_mod <- glm(Marijuana ~ Gender + Age + Race + Employment, data = m4, family = "binomial")
e_hat <- ps_mod$fitted.values
dat_1 <- transform(m4, Marijuana = 1)
dat_0 <- transform(m4, Marijuana = 0)
y_hat_1 <- predict(out_mod, newdata = dat_1, type = "response") 
y_hat_0 <- predict(out_mod, newdata = dat_0, type = "response")
boxplot(y_hat_1, y_hat_0)
DR_1 <- (e_hat^-1)*y*trt - (e_hat^-1)*(y_hat_1*(trt - e_hat))
DR_0 <- (1 - e_hat)^-1*(y*(1 - trt)) + (1 - e_hat)^-1*(y_hat_0*(trt - e_hat))
paste("Estimated treatment effect (ATE): ",mean(DR_1) - mean(DR_0))
```

