---
title: "Observational Studies Midterm Project Notebook"
output:
  html_document:
    df_print: paged
---

# Observational Studies Midterm Project

Read in data.
```{r}
rm(list = ls())
library(tidyverse)
library(gridExtra)
library(grid)
library(knitr)
library(xtable)
library(optmatch)
library(RItools)
library(scales)
library(RColorBrewer)
library(plyr)
df_2017 = read.table(file = 'NSDUH_2017_Tab.tsv', sep = '\t', header = TRUE)
```

Write functions to extract certain types of data. Use `QUESTID2` as an ID for each dataset.

(a) Demographics Data 
```{r}
read_dem = function(df){
  dem_df = df %>% select("ID" = QUESTID2, #ID 
                   "Gender" = IRSEX, # Gender (M/F)
                   "Age" = CATAG7, # Age category (1-7, 12-13, 14-15, 16-17, 18-20, 21-25, 26-34, 35+)
                   "Marital" = IRMARIT, # Marital Status (1/2/3/4/99 - Married, widowed, divorced or separated, never been married, underage)
                   "Health" = HEALTH, # Overall health (1-5, Excellent, Very good, good, fair, poor)
                   "Education" = EDUSCHGRD2, # Education (1-11, Fifth grade of less, sixth - 12th, freshmen/13th year, sophomore or junior / 14th or 15th year, senior / 16th year or higher)
                   "Race" = NEWRACE2, # Race (1-7, White/African American/ Native American/ Native Hawaiian or other Pacific Island/ Asian / More than one race / Hispanic)
                   "College" = COLLENRLST, # College enrollment (./1/2/3/4/5, Unkonwn, Full time, Part time, Not enrolled , Enrolled with no info, Outside of age 18-22)
                   "Employment" = IRWRKSTAT, # Employment Status (1/2/3/4/99, Full time, part time, unemployed, other, underage)
                   "PopDen" = PDEN10 # Population Density (1/2/3, !mil or more, 1mil or less, unknown)
  ) %>%
    mutate(Gender = recode(Gender, `1` = "M", `2` = "F", .default = NULL),
           Age = recode_factor(Age, `1` = "12-13", `2` = "14-15", `3` = "16-17", 
                               `4` = "18-20", `5` = "21-25", `6` = "26-34", 
                               `7` = "35+", .ordered = TRUE, .default = NULL),
           Marital = recode(Marital, `1` = "Married", `2` = "Widowed", `3` = "Divorced/Separated",
                            `4` = "Unmarried", `99` = "Underage"),
           Health = recode_factor(Health, `5` = "Poor", `4` = "Fair", `3` = "Good",
                                  `2` = "Very Good", `1` = "Excellent", 
                                    .default = "Unknown", .ordered = TRUE),
           Education = recode_factor(Education, `1` = "Fifth or less", `2` = "Sixth", `3` = "Seventh",
                                     `4` = "Eighth", `5` = "Ninth", `6` = "Tenth", `7` = "Eleventh",
                                     `8` = "Twelfth", `9` = "Freshman", `10` = "Soph/Junior", 
                                     `11` = "Senior or higher", .default = NULL, .ordered = TRUE),
           Race = recode(Race, `1` = "White (Nonhispanic)", `2` = "Black", `3` = "Native American",
                         `4` = "Native Hawaiian", `5` = "Asian", `6` = "Multiracial",
                         `7` = "Hispanic", .default = NULL),
           College = recode(College, `1` = "Full time", 
                            `2` = "Part time", `3` = "Not enrolled",
                            `4` = "Full time", `5` = "Underage", 
                            .default = "Not enrolled", .missing = "Unknown"),
           Employment = recode(Employment, `1` = "Full time", `2` = "Part time", `3` = "Unemployed",
                               `4` = "Other", `99` = "Underage", .default = NULL),
           PopDen = recode(PopDen, `1` = "Large", `2` = "Small", `3` = "Unknown")
           )
  return(dem_df)
}
```

(b) Drug Data
Get data on: ever done it? age when first done? last time since it was done? how many times in the past 30 days?
Coding for numbers: 
- 1/2 - Yes/No
- 1/2/91 - Yes/No/never used umbrella drug
- a/b/991 - age or never used
- a/b/991/993 - age or never used or never used in past 12 months
- 1/2/3/91 - within last 30 days, within last year, more than a year, never used
- 1/2/3/4/91 - Within last 30 days, within last year, within 3 years, more than 3 years, never used
- 1-30/91/93 - days or never used or never used in past 30 days 
```{r}
# Read in drug info
read_drugs = function(df){
  drug_df = df %>% select("ID" = QUESTID2, # ID
                          "Cigarette" = CIGEVER, # Ever smoked a cigarette (1/2)
                          "CigaretteFirstAge" = CIGTRY, # Age when first smoked a cigarette (a-b/991)
                          "CigaretteLastUse" = CIGREC, # Time since last smoked (1/2/3/4/91)
                          "CigaretteLast30Days" = CIG30USE, # How many days smoked cigarette in past 30 days (a-b/91/93)
     
                          "Marijuana" = MJEVER, # Ever used marijuana (1/2)
                          "MarijuanaFirstAge" = MJAGE, # Age when first used mj/hashish (a-b/991)
                          "MarijuanaLastUse" = MJREC, # Time since last used mj or hashish (1/2/3/91)
                          "MarijuanaLastYear" = MRDAYPYR # Number of days used mj/hashish in past 12 months (0-365/991/993) 

  ) %>%
    mutate_at(vars(Cigarette, Marijuana),
              ~recode(., `1` = "Yes", `2` = "No", .default = "Unknown")) %>% 
    mutate_at(vars(CigaretteFirstAge, MarijuanaFirstAge), 
              list(~ ifelse(. %in% c(985, 991, 994, 997, 998), "Unknown", .))) %>%
    mutate_at(vars(CigaretteLastUse, MarijuanaLastUse), 
          ~recode_factor(., .default = "Never Used",`1` = "Past 30 Days", 
                         `2` = "Past Year", `3` = "Past 3 Years", 
                         `4` = "More than 3 Years", .ordered = TRUE)) %>%
    mutate(CigaretteLast30Days = ifelse(CigaretteLast30Days %in% c(91, 93, 94, 97, 98), 
                                        "Unknown", CigaretteLast30Days)) %>%
    mutate_at(vars(MarijuanaLastYear), 
              list(~ ifelse(. %in% c(seq(1, 365, 1), 991), ., "Unknown")))
  return(drug_df)
}
```

(c) Criminal Data
Coding for numbers: 
- 1/2 - Yes/No
- 0/1/2/3 - 0 times, 1 time, 2 times, 3 or more times
```{r}
# Criminal info
read_crim = function(df){
  crim_df = df %>% select("ID" = QUESTID2, # ID
                          "TimesArrested" = NOBOOKY2, # Number of times arrested and booked for breaking the law in the past 12 months (0/1/2/3+)
                          "MVTheft" = BKMVTHFT, # Ever arrested and booked for motor vehicle theft in past 12 months (1/2)
                          "Theft" = BKLARCNY, # Ever arrested and booked for larceny/theft in past 12 months (1/2)
                          "Burglary" = BKBURGL, # Ever arrested and booked for burglary or breaking and entering in past 12 months (1/2)
                          "ViolentOffense" = BKSRVIOL, # Ever arrested and booked for serious violent offenses in past 12 monthsm including aggravated assault, forcible rape, murder, homicide or nonnegligant manslaughter (1/2)
                          "Robbery" = BKROB, # Ever arrested and booked for robbery in past 12 months (1/2)
                          "Arson" = BKARSON, # Ever arrested and booked for arson in past 12 months (1/2)
                          "DUI" = BKDRVINF, # Ever arrested and booked for DUI (driving under influence of alcohol of drugs) in past 12 months (1/2)
                          "PossSale" = BKDRUG, # Ever arrested and booked for possession/manufacture or sale of drugs (1/2)
                          "SexualOffense" = BKSEXNR, # Ever arrested and booked for prostitution or commercialized sex or any other sexual offense not including rape in past 12 months (1/2)
                          "Fraud" = BKFRAUD # Ever arrested and booked for fraud, possessing stolen goods or vandalism in past 12 months (1/2)
  ) %>% 
    mutate(TimesArrested = ifelse(TimesArrested <= 3, TimesArrested, NA),
           TimesArrested = recode_factor(TimesArrested, `0` = "None", `1` = "One Time",
                                  `2` = "Two Time", `3` = "Three or more times",
                                  .ordered = TRUE)) %>%
    mutate_at(vars(MVTheft, Theft, Burglary, ViolentOffense, Robbery, Arson, 
                   DUI, PossSale, SexualOffense, Fraud), 
              list(~ ifelse(. < 3, ., NA))) %>%
    mutate_at(vars(MVTheft, Theft, Burglary, ViolentOffense, Robbery, Arson, 
                   DUI, PossSale, SexualOffense, Fraud), 
              list(~ ifelse(. == 1, "Yes", "No")))
  return(crim_df)
}
```

(d) Emotional Stress Indicators
Coding for numbers:
- 1/2/3/4/5 - all of the time, most of the time, some of the time, a little of the time, none of the time
- 1/2 - Yes/No
- 1/2/3/4 - No difficulty, mild difficulty, moderate difficulty, severe difficulty
- 1-52 - number of weeks 
- 1/2/3 - Every day, most days, only one or two days a week
- 0-365 - number of days 
```{r}
read_estress = function(df){
  mh_df = df %>% select("ID" = QUESTID2, # ID
                        "Nervous30" = DSTNRV30, # How often did you feel nervous in past 30 days (1/2/3/4/5)
                        "Hopeless30" = DSTHOP30, # How often did you feel hopeless in past 30 days (1/2/3/4/5)
                        "Worthless30" = DSTNGD30, # How often did you feel down on yourself, no good or worthless in the past days (1/2/3/4/5)
                        
                        "WorstStressedMonth" = DSTWORST, # Was there a month in the past 12 month months where you felt more depressed/ anxious or emotionally stressed than the last 30 days (1/2)
                        "NervousWorst" = DSTNRV12, # In that one month, how often did you feel nervous (1/2/3/4/5)
                        "HopelessWorst" = DSTHOP12, # In that one month, how often did you feel hopeless (1/2/3/4/5)
                        "WorthlessWorst" = DSTNGD12 # In that one month, how often did you feel down on yourself, no good or worthless (1/2/3/4/5)
  ) %>%
    mutate_at(vars(Nervous30, Hopeless30, Worthless30,
                   NervousWorst, HopelessWorst, WorthlessWorst),
              ~recode_factor(., 
                             `5` = "None of the time", 
                             `4` = "A little of the time",
                             `3` = "Some of the time", 
                             `2` = "Most of the time",
                             `1` = "All of the time", 
                             .default = NULL, 
                             .ordered = TRUE)) 
  return(mh_df)
}
```


Read in all data.
```{r, warning = FALSE}
dem_2017 = read_dem(df_2017)
drugs_2017 = read_drugs(df_2017)
crim_2017 = read_crim(df_2017)
es_2017 = read_estress(df_2017)
```


```{r}
drug_plot = function(df1, df2, drug, ymin, ymax){
  
  num_peops = as.data.frame(rbind(table(df1[[drug]]), table(df2[[drug]])))
  num_peops$No = -num_peops$No
  plot = num_peops %>%
    gather(yesno, num_people, Yes, No) %>%
    ggplot(aes(x = year, y = num_people, fill = yesno)) + 
    geom_segment(aes(x = year, xend = year, y = 0, yend = num_people, color = yesno), size = 1.25) + 
    geom_point( aes(x=year, y=0), size=1, show.legend = FALSE) +
    geom_point( aes(x=year, y=num_people), size=1.15, show.legend = FALSE) +
    scale_color_discrete(name = NULL, labels = c("Not Taken", "Taken")) +
    scale_y_continuous(breaks = c(ymin, 0, ymax), labels = c(abs(ymin), 0, ymax)) + 
    labs(x = "Year", y = "Number of People") + 
    ggtitle(drug) + coord_flip()
  return(plot) 
}
```

Function to put plots together with shared legends
```{r}
grid_arrange_shared_legend <- function(..., ncol = length(list(...)), nrow = 1, position = c("bottom", "right")) {
  plots <- list(...)
  position <- match.arg(position)
  g <- ggplotGrob(plots[[1]] + theme(legend.position = position))$grobs
  legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
  lheight <- sum(legend$height)
  lwidth <- sum(legend$width)
  gl <- lapply(plots, function(x) x + theme(legend.position="none"))
  gl <- c(gl, ncol = ncol, nrow = nrow)
  combined <- switch(position,
                     "bottom" = arrangeGrob(do.call(arrangeGrob, gl),
                                            legend,
                                            ncol = 1,
                                            heights = unit.c(unit(1, "npc") - lheight, lheight)),
                     "right" = arrangeGrob(do.call(arrangeGrob, gl),
                                           legend,
                                           ncol = 2,
                                           widths = unit.c(unit(1, "npc") - lwidth, lwidth)))
  grid.newpage()
  grid.draw(combined)
  # return gtable invisibly
  invisible(combined)
}
```

Total Number of Drug Users
```{r}
x = drugs_2017 %>% select(Cigarette) %>% count()
colnames(x) = c("Yes/No", "Cigarette")
y = drugs_2017 %>% filter(Marijuana %in% c("Yes", "No")) %>% select(Marijuana) %>% count()
colnames(y) = c("Yes/No", "Marijuana")
total_df = merge(x,y)
total_df
print(xtable(total_df, type = "latex", caption = "Figure 1: Usage of Cigarette and Marijuana"), file = "cig_mar_users.tex")
```

Covariate Plots - Gender, Age, Race, Employment vs Cigarette, Marijuana
```{r}
cig_cov = inner_join(dem_2017, drugs_2017, by = "ID")
```

Plot of Covariates for Cigarettes
```{r}
j1 = cig_cov %>% select(Gender, Cigarette) %>% count() %>%
  mutate("perc" = ifelse(Cigarette == "Yes", freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Gender, y = perc, fill = Cigarette)) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.4, 0.4),
                     breaks = seq(-0.4, 0.4, by = 0.2), 
                     labels = c(0.4, 0.2, 0, 0.2, 0.4)) + 
  labs(x = "Gender", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Gender in 2017") + 
  scale_fill_manual(values = brewer.pal(3, "Accent")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

j2 = cig_cov %>% select(Age, Cigarette) %>% count() %>%
  mutate("perc" = ifelse(Cigarette == "Yes", freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Age, y = perc, fill = Cigarette)) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.4, 0.4),
                     breaks = seq(-0.4, 0.4, by = 0.2),
                     labels = c(0.4, 0.2, 0, 0.2, 0.4)) + 
  labs(x = "Age", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Age in 2017") + 
  scale_fill_manual(values = brewer.pal(3, "Accent")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

j3 = cig_cov %>% select(Race, Cigarette) %>% count() %>%
  mutate("perc" = ifelse(Cigarette == "Yes", freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Race, y = perc, fill = Cigarette)) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.4, 0.4),
                     breaks = seq(-0.4, 0.4, by = 0.2), 
                     labels = c(0.4, 0.2, 0, 0.2, 0.4)) + 
  labs(x = "Race", y = "percentage", title = "Percentage Distribution", 
       subtitle = " by Race in 2017") + 
  scale_fill_manual(values = brewer.pal(2, "Accent")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

j4 = cig_cov %>% select(Employment, Cigarette) %>% count() %>%
  mutate("perc" = ifelse(Cigarette == "Yes", freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Employment, y = perc, fill = Cigarette)) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.4, 0.4), 
                     breaks = seq(-0.4, 0.4, by = 0.2), 
                     labels = c(0.4, 0.2, 0, 0.2, 0.4)) + 
  labs(x = "Employment", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Employment in 2017") + 
  scale_fill_manual(values = brewer.pal(2, "Accent")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

j_all = grid_arrange_shared_legend(j1, j2, j3, j4, nrow = 2, ncol = 2)
ggsave("CigaretteDemographics.png", j_all, height = 7, width = 7, units = "in")
```

Plot of Covariates for Marijuana
```{r}
k1 = cig_cov %>% select(Gender, Marijuana) %>% filter(Marijuana %in% c("Yes", "No")) %>% count() %>%
  mutate("perc" = ifelse(Marijuana == "Yes", freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Gender, y = perc, fill = Marijuana)) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.4, 0.4),
                     breaks = seq(-0.4, 0.4, by = 0.2), 
                     labels = c(0.4, 0.2, 0, 0.2, 0.4)) + 
  labs(x = "Gender", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Gender in 2017") + 
  scale_fill_manual(values = brewer.pal(3, "Accent")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

k2 = cig_cov %>% select(Age, Marijuana) %>% filter(Marijuana %in% c("Yes", "No")) %>% count() %>%
  mutate("perc" = ifelse(Marijuana == "Yes", freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Age, y = perc, fill = Marijuana)) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.4, 0.4),
                     breaks = seq(-0.4, 0.4, by = 0.2),
                     labels = c(0.4, 0.2, 0, 0.2, 0.4)) + 
  labs(x = "Age", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Age in 2017") + 
  scale_fill_manual(values = brewer.pal(3, "Accent")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

k3 = cig_cov %>% select(Race, Marijuana) %>% filter(Marijuana %in% c("Yes", "No")) %>% count() %>%
  mutate("perc" = ifelse(Marijuana == "Yes", freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Race, y = perc, fill = Marijuana)) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.4, 0.4),
                     breaks = seq(-0.4, 0.4, by = 0.2), 
                     labels = c(0.4, 0.2, 0, 0.2, 0.4)) + 
  labs(x = "Race", y = "percentage", title = "Percentage Distribution", 
       subtitle = " by Race in 2017") + 
  scale_fill_manual(values = brewer.pal(2, "Accent")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

k4 = cig_cov %>% select(Employment, Marijuana) %>% filter(Marijuana %in% c("Yes", "No")) %>% count() %>%
  mutate("perc" = ifelse(Marijuana == "Yes", freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Employment, y = perc, fill = Marijuana)) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.4, 0.4), 
                     breaks = seq(-0.4, 0.4, by = 0.2), 
                     labels = c(0.4, 0.2, 0, 0.2, 0.4)) + 
  labs(x = "Employment", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Employment in 2017") + 
  scale_fill_manual(values = brewer.pal(2, "Accent")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

k_all = grid_arrange_shared_legend(k1, k2, k3, k4, nrow = 2, ncol = 2)
ggsave("MarijuanaDemographics.png", k_all, height = 7, width = 7, units = "in")
```


Join tables for analysis
```{r}
# Join drugs and crimes and extract relevant variables.
# Remove rows of no response
drug_crimes_vars = function(dem_df, drug_df, crime_df){
  return(inner_join(dem_df, drug_df, by = "ID") %>% left_join(., crime_df, by="ID"))
}

drug_crime_spec = function(drug){
  df1 = drug_crimes_vars(dem_2017, drugs_2017 %>% select(ID, drug), crim_2017) %>% mutate(year = 2017)
  df_overall = df1 %>% subset(select = -ID)
  return(df_overall)
}

drug_es_vars = function(dem_df, drug_df, mh_df){
  return(inner_join(dem_df, drug_df, by = "ID") %>% left_join(., mh_df, by="ID"))
}

drug_es_spec = function(drug){
  df1 = drug_es_vars(dem_2017, drugs_2017 %>% select(ID, drug), es_2017) %>% mutate(year = 2017)
  df_overall = df1 %>% subset(select = -ID)
  return(df_overall)
}
```

Create vectors of relevant variables
```{r}
crimes = c("Theft", "Arson", "Burglary")
worst_feelings = c("NervousWorst", "HopelessWorst", "WorthlessWorst")
```

Create new table - Crime/Cigarette
```{r}
crime_cig_2017 = na.omit(drug_crime_spec("Cigarette") %>% 
                      filter(Cigarette %in% c("Yes", "No")))
crime_cig_2017$Cigarette = ifelse(crime_cig_2017$Cigarette == "Yes", 1, 0)
crime_cig_2017[crimes] = crime_cig_2017[crimes] %>% 
  mutate_all(., ~recode(., `Yes` = 1, `No` = 0))
crime_cig_2017 = crime_cig_2017 %>% select(Gender, Age, Race, Employment, crimes, Cigarette) %>% 
  mutate("crimes" = ifelse((Theft == 1) | (Arson == 1) | (Burglary == 1), 1, 0)) %>% subset(select = -c(Theft, Arson, Burglary))
```

Create new table - Crime/Marijuana
```{r}
crime_mar_2017 = na.omit(drug_crime_spec("Marijuana") %>%
                           filter(Marijuana %in% c("Yes", "No")))
crime_mar_2017$Marijuana = ifelse(crime_mar_2017$Marijuana == "Yes", 1, 0)
crime_mar_2017[crimes] = crime_mar_2017[crimes] %>% 
  mutate_all(., ~recode(., `Yes` = 1, `No` = 0))
crime_mar_2017 = crime_mar_2017 %>% select(Gender, Age, Race, Employment, crimes, Marijuana) %>% 
  mutate("crimes" = ifelse((Theft == 1) | (Arson == 1) | (Burglary == 1), 1, 0)) %>% subset(select = -c(Theft, Arson, Burglary))
```

Create new table - Emotional Stress/Cigarette
```{r}
es_cig_2017 = na.omit(drug_es_spec("Cigarette") %>% 
                      filter(Cigarette %in% c("Yes", "No")))
es_cig_2017$Cigarette = ifelse(es_cig_2017$Cigarette == "Yes", 1, 0)
es_cig_2017[worst_feelings] = es_cig_2017[worst_feelings] %>% 
  mutate_all(. ~recode(., "None of the time" = 0, 
                       "A little of the time" = 1, "Some of the time" = 3, 
                       "Most of the time" = 4, "All of the time" = 5))
es_cig_2017 = es_cig_2017 %>% select(Gender, Age, Race, Employment, worst_feelings, Cigarette) %>% 
  mutate("feelings" = ifelse((NervousWorst > 2) | (HopelessWorst > 2) | (WorthlessWorst > 2), 1, 0)) %>%  
  subset(select = -c(NervousWorst, HopelessWorst, WorthlessWorst))
```

Create new table - Emotional Stress/Marijuana
```{r}
es_mar_2017 = na.omit(drug_es_spec("Marijuana") %>%
                           filter(Marijuana %in% c("Yes", "No")))
es_mar_2017$Marijuana = ifelse(es_mar_2017$Marijuana == "Yes", 1, 0)
es_mar_2017[worst_feelings] = es_mar_2017[worst_feelings] %>% 
  mutate_all(. ~recode(., "None of the time" = 0, 
                       "A little of the time" = 1, "Some of the time" = 3, 
                       "Most of the time" = 4, "All of the time" = 5))
es_mar_2017 = es_mar_2017 %>% select(Gender, Age, Race, Employment, worst_feelings, Marijuana) %>% 
  mutate("feelings" = ifelse((NervousWorst > 2) | (HopelessWorst > 2) | (WorthlessWorst > 2), 1, 0)) %>% 
  subset(select = -c(NervousWorst, HopelessWorst, WorthlessWorst))
```

## Modeling 

### Model 1: Cigarette / Crimes
```{r}
model1 = glm(Cigarette ~ Gender + Age + Race + Employment + crimes, 
             data = crime_cig_2017, family = "binomial")
a1 = match_on(model1, data = crime_cig_2017)
b1 = pairmatch(a1, data = crime_cig_2017)
c1 = glm(crimes ~ Age + Gender + Race + Employment + Cigarette, 
        data = na.omit(crime_cig_2017[b1,]), family = "binomial")
100 * (exp(as.numeric(tail(summary(c1)$coef[,1], 1))) - 1)
```

Model 1 Covariates after matching
```{r}
m1 = na.omit(crime_cig_2017[b1,]) %>% select(Gender, Cigarette) %>% count() %>%
  mutate("perc" = ifelse(Cigarette == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Gender, y = perc, fill = as.factor(Cigarette))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Gender", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Gender in 2017") + 
  scale_fill_manual(name = "Cigarette", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m2 = na.omit(crime_cig_2017[b1,]) %>% select(Age, Cigarette) %>% count() %>%
  mutate("perc" = ifelse(Cigarette == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Age, y = perc, fill = as.factor(Cigarette))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Age", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Age in 2017") + 
  scale_fill_manual(name = "Cigarette", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m3 = na.omit(crime_cig_2017[b1,]) %>% select(Race, Cigarette) %>% count() %>%
  mutate("perc" = ifelse(Cigarette == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Race, y = perc, fill = as.factor(Cigarette))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Race", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Race in 2017") + 
  scale_fill_manual(name = "Cigarette", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m4 = na.omit(crime_cig_2017[b1,]) %>% select(Employment, Cigarette) %>% count() %>%
  mutate("perc" = ifelse(Cigarette == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Employment, y = perc, fill = as.factor(Cigarette))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Employment", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Employment in 2017") + 
  scale_fill_manual(name = "Cigarette", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m_grid = grid_arrange_shared_legend(m1, m2, m3, m4, ncol = 2, nrow = 2)
ggsave("model1.png", m_grid, height = 7, width = 7, units = "in")
```

### Model 2: Marijuana / Crimes
```{r}
model2 = glm(Marijuana ~ Gender + Age + Race + Employment + crimes, 
             data = crime_mar_2017, family = "binomial")
a2 = match_on(model2, data = crime_mar_2017)
b2 = pairmatch(a2, data = crime_mar_2017)
c2 = glm(crimes ~ Age + Gender + Race + Employment + Marijuana, 
      data = na.omit(crime_mar_2017[b2,]), family = "binomial")
100 * (exp(as.numeric(tail(summary(c2)$coef[,1], 1))) - 1)
```

Model 2 Covariates after Matching
```{r}
m1 = na.omit(crime_mar_2017[b2,]) %>% select(Gender, Marijuana) %>% count() %>%
  mutate("perc" = ifelse(Marijuana == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Gender, y = perc, fill = as.factor(Marijuana))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Gender", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Gender in 2017") + 
  scale_fill_manual(name = "Marijuana", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m2 = na.omit(crime_mar_2017[b2,]) %>% select(Age, Marijuana) %>% count() %>%
  mutate("perc" = ifelse(Marijuana == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Age, y = perc, fill = as.factor(Marijuana))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Age", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Age in 2017") + 
  scale_fill_manual(name = "Marijuana", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m3 = na.omit(crime_mar_2017[b2,]) %>% select(Race, Marijuana) %>% count() %>%
  mutate("perc" = ifelse(Marijuana == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Race, y = perc, fill = as.factor(Marijuana))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Race", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Race in 2017") + 
  scale_fill_manual(name = "Marijuana", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m4 = na.omit(crime_mar_2017[b2,]) %>% select(Employment, Marijuana) %>% count() %>%
  mutate("perc" = ifelse(Marijuana == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Employment, y = perc, fill = as.factor(Marijuana))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Employment", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Employment in 2017") + 
  scale_fill_manual(name = "Marijuana", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m_grid = grid_arrange_shared_legend(m1, m2, m3, m4, ncol = 2, nrow = 2)
ggsave("model2.png", m_grid, height = 7, width = 7, units = "in")
```

### Model 3: Cigarette / Emotional Stress
```{r}
model3 = glm(Cigarette ~ Gender + Age + Race + Employment + feelings, 
             data = es_cig_2017, family = "binomial")
a3 = match_on(model3, data = es_cig_2017)
b3 = pairmatch(a3, data = es_cig_2017)
c3 = glm(feelings ~ Age + Gender + Race + Employment + Cigarette, 
        data = na.omit(es_cig_2017[b3,]), family = "binomial")
100 * (exp(as.numeric(tail(summary(c3)$coef[,1], 1))) - 1)
```

Model 3 Covariates after Matching
```{r}
m1 = na.omit(es_cig_2017[b3,]) %>% select(Gender, Cigarette) %>% count() %>%
  mutate("perc" = ifelse(Cigarette == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Gender, y = perc, fill = as.factor(Cigarette))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Gender", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Gender in 2017") + 
  scale_fill_manual(name = "Cigarette", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m2 = na.omit(es_cig_2017[b3,]) %>% select(Age, Cigarette) %>% count() %>%
  mutate("perc" = ifelse(Cigarette == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Age, y = perc, fill = as.factor(Cigarette))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Age", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Age in 2017") + 
  scale_fill_manual(name = "Cigarette", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m3 = na.omit(es_cig_2017[b3,]) %>% select(Race, Cigarette) %>% count() %>%
  mutate("perc" = ifelse(Cigarette == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Race, y = perc, fill = as.factor(Cigarette))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Race", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Race in 2017") + 
  scale_fill_manual(name = "Cigarette", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m4 = na.omit(es_cig_2017[b3,]) %>% select(Employment, Cigarette) %>% count() %>%
  mutate("perc" = ifelse(Cigarette == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Employment, y = perc, fill = as.factor(Cigarette))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Employment", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Employment in 2017") + 
  scale_fill_manual(name = "Cigarette", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m_grid = grid_arrange_shared_legend(m1, m2, m3, m4, ncol = 2, nrow = 2)
ggsave("model3.png", m_grid, height = 7, width = 7, units = "in")
```

### Model 4: Marijuana / Emotional Stress
```{r}
model4 = glm(Marijuana ~ Gender + Age + Race + Employment + feelings, 
             data = es_mar_2017, family = "binomial")
a4 = match_on(model4, data = es_mar_2017)
b4 = pairmatch(a4, data = es_mar_2017)
c4 = glm(feelings ~ Gender + Age + Race + Employment + Marijuana, 
        data = na.omit(es_mar_2017[b4,]), family = "binomial")
100 * (exp(as.numeric(tail(summary(c4)$coef[,1], 1))) - 1)
```

Model 4 Covariates after Matching
```{r}
m1 = na.omit(es_mar_2017[b4,]) %>% select(Gender, Marijuana) %>% count() %>%
  mutate("perc" = ifelse(Marijuana == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Gender, y = perc, fill = as.factor(Marijuana))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Gender", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Gender in 2017") + 
  scale_fill_manual(name = "Cigarette", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m2 = na.omit(es_mar_2017[b4,]) %>% select(Age, Marijuana) %>% count() %>%
  mutate("perc" = ifelse(Marijuana == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Age, y = perc, fill = as.factor(Marijuana))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Age", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Age in 2017") + 
  scale_fill_manual(name = "Cigarette", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m3 = na.omit(es_mar_2017[b4,]) %>% select(Race, Marijuana) %>% count() %>%
  mutate("perc" = ifelse(Marijuana == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Race, y = perc, fill = as.factor(Marijuana))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Race", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Race in 2017") + 
  scale_fill_manual(name = "Cigarette", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m4 = na.omit(es_mar_2017[b4,]) %>% select(Employment, Marijuana) %>% count() %>%
  mutate("perc" = ifelse(Marijuana == 1, freq / sum(freq), -1 * freq / sum(freq))) %>% 
  ggplot(aes(x = Employment, y = perc, fill = as.factor(Marijuana))) + geom_bar(stat = "identity") + 
  scale_y_continuous(limits = c(-0.6, 0.6),
                     breaks = seq(-0.6, 0.6, by = 0.3), 
                     labels = c(0.6, 0.3, 0, 0.3, 0.6)) + 
  labs(x = "Employment", y = "percentage", 
       title = "Percentage Distribution", 
       subtitle = " by Employment in 2017") + 
  scale_fill_manual(name = "Cigarette", values = brewer.pal(3, "Set2")) + 
  coord_flip() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")

m_grid = grid_arrange_shared_legend(m1, m2, m3, m4, ncol = 2, nrow = 2)
ggsave("model4.png", m_grid, height = 7, width = 7, units = "in")
```

Create summary table for models
```{r}
coefs = round(c(tail(summary(c1)$coef,1)[1], 
                  tail(summary(c2)$coef,1)[1],
                  tail(summary(c3)$coef,1)[1],
                  tail(summary(c4)$coef,1)[1]), 3)

pvals = round(c(tail(summary(c1)$coef,1)[4], 
                  tail(summary(c2)$coef,1)[4],
                  tail(summary(c3)$coef,1)[4],
                  tail(summary(c4)$coef,1)[4]), 3)

model_name = c("Cigarette/Crimes", "Marijuana/Crimes",
               "Cigarette/Stress", "Marijuana/Stress")

matched_num = c(stratumStructure(b1)[2],
                stratumStructure(b2)[2],
                stratumStructure(b3)[2],
                stratumStructure(b4)[2])
unmatched_num = c(stratumStructure(b1)[1],
                  stratumStructure(b2)[1],
                  stratumStructure(b3)[1],
                  stratumStructure(b4)[1])

model_df = data.frame("Model" = model_name, 
                      "Coefficient of Crime" = coefs,
                      "Number of Matches" = matched_num, 
                      "Number of Nonmatches" = unmatched_num,
                      "p value" = pvals)
model_df$Interpretation = 100* (exp(model_df$Coefficient.of.Crime) - 1)
model_df
print(xtable(model_df, type = "latex", 
             caption = "Table 2: Propensity Score Matches and Logistic Regression Model Coefficient"), 
      file = "coefs.tex")
```
